\paragraph{NeurIPS 2021 Paper Checklist.}

\begin{itemize}
    \item 
    Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    
    \answerYes{See abstract, Section 1, and Section 2 for the discussions on our main claims.}
    
    \item
    Have you read the ethics review guidelines and ensured that your paper conforms to them?
    
    \answerYes{}
    
    \item
    Did you discuss any potential negative societal impacts of your work?
    
    \answerTODO{}
    
    \item
    Did you describe the limitations of your work?
    
    \answerYes{See Section 6 for our discussion on the limitations.}
    
    \item
    Did you state the full set of assumptions of all theoretical results?
    
    \answerYes{See Section 3 and Section 4.}
    
    \item
    Did you include complete proofs of all theoretical results?
    
    \answerYes{Complete proofs are included in Appendix A and Appendix B.}
    
    \item
    Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    
    \answerTODO{}
    
    \item
    Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    
    \answerYes{See Section 5 for the simulations.}
    
    \item
    Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    
    \answerYes{See Figure 2 and Figure 3.}
    
    \item
    Did you include the amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    
    \answerTODO{}
    
    \item
    If your work uses existing assets, did you cite the creators?
    
    \answerTODO{}
    
    \item
    Did you mention the license of the assets?
    
    \answerTODO{}
    
    \item
    Did you include any new assets either in the supplemental material or as a URL?
    
    \answerTODO{}
    
    \item
    Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    
    \answerNA{Only external data used are MNIST data.}
    
    \item
    Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    
    \answerNA{Only external data used are MNIST data.}
\end{itemize}